---
title: "Project Uccle"
author: "Marco Ciotola"
output:
  pdf_document:
    keep_tex: yes
    fig_width: 6.5
    fig_height: 4
#  html_document:
#    mathjax: local   
#    self_contained: no
#    fig_width: 6.5
#    fig_height: 4
---

```{r tentatives, eval=FALSE, child='chapter1.Rmd', include=FALSE}
```

```{r set uccs, include=FALSE}
# UCCS
domain <- c(-350000,350000); pattern <- "UCCS.*\\.csv"; name <- "UCCS"; description <- "Uccle, Bruxelles"; proj_file <- 'proj.uccs.RData'
```

```{r project, child='project.Rmd'}
```

## Arima

```{r arima}
to_test <- list(
  #c(0,0,0,0,0,1),  # BAD, Box-Pierce = 0
  #c(0,0,0,0,0,2),  # BAD, Box-Pierce = 0
  #c(0,0,0,0,1,1),  # nice, Box-Pierce = 0
  c(0,0,0,0,1,2),
  #c(0,0,0,0,1,3),  # CAN'T BE DONE: 3*168 = 504 > 350 (limit of arima on r)
  #c(0,1,0,0,0,1),  # BAD, Box-Pierce = 0
  c(0,1,0,0,1,1),
  c(0,1,0,0,1,2),
  c(2,0,0,0,1,0),  # hint from auto.arima D=1
  c(5,0,0,0,1,0),  # hint from auto.arima D=1
  c(3,1,1,0,1,0),  # hint from auto.arima D=1 d=1
  c(3,1,1,0,1,1),  # after hint then ACF + PACF
  c(4,1,2,0,1,0),  # hint from auto.arima d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10
  c(4,1,2,0,1,1)  # after hint then ACF + PACF
)
if (!exists("models")){
  models <- list()
}
for (model in to_test) {
  # model <- c(4,1,2,0,1,0)
  fit.name <- paste0("fit", paste0(model, collapse = ""))
  forecast.name <- paste0("forecast", paste0(model, collapse = ""))
  print("SARIMA(" %s+% paste0(model[1:3], collapse = ",") %s+% ")(" %s+% paste0(model[4:6], collapse = ",") %s+% ")[" %s+% frequency(seis.h.train) %s+% "]")
  if (!(fit.name %in% names(models))) {
    models[[fit.name]] <- Arima(seis.h.train, model[1:3], model[4:6])
  }
  if (!(forecast.name %in% names(models))) {
    models[[forecast.name]] <- forecast(models[[fit.name]], h=168*18)
  }
}
#models[["fitauto"]] <- auto.arima(seis.h.train, D=1, approx=F, trace = T)
#models[["forecastauto"]] <- forecast(models[["fitauto"]], h=168*18)  
#models[["fit311010"]] <- models[["fitauto"]]
#models[["fit412010"]] <- fit
#models[["forecast412010"]] <- forecast(models[["fit412010"]], h=168*18)

# 000012 better than 000011, Box-Pierce = 0
plotting.model <- models[["fit000012"]]
plotting.forecast <- models[["forecast000012"]]
# 010011 somehow better somehow worst than 000012, Box-Pierce = 0
plotting.model <- models[["fit010011"]]
plotting.forecast <- models[["forecast010011"]]
# 010012 , Box-Pierce = 
plotting.model <- models[["fit010012"]]
plotting.forecast <- models[["forecast010012"]]
# 311010 , Box-Pierce = 
plotting.model <- models[["fit311010"]]
plotting.forecast <- models[["forecast311010"]]
# 311011 , Box-Pierce = 
plotting.model <- models[["fit311011"]]
plotting.forecast <- models[["forecast311011"]]

checkresiduals(plotting.model)
residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
plot_grid(residuals.acf, residuals.pacf, nrow=2)
plot_forecast(plotting.forecast) + 
  autolayer(seis.h.test, series="Original")
plots <- show_overlap_and_diff(
  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
  base.desc = "original series",
  compare.desc = "reconstructed series",
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="% over ts span",
  diff.ylab.sec.accuracy = .01
)
plot_grid(plots[[1]], plots[[2]], nrow = 2)
accuracy(plotting.forecast, seis.h.test)




plotting.model <- models[["fit010012"]]
plotting.forecast <- models[["forecast010012"]]
checkresiduals(plotting.model)
residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
plot_grid(residuals.acf, residuals.pacf, nrow=2)
plot_forecast(plotting.forecast) + 
  autolayer(seis.h.test, series="Original")
plots <- show_overlap_and_diff(
  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
  base.desc = "original series",
  compare.desc = "reconstructed series",
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="% over ts span",
  diff.ylab.sec.accuracy = .01
)
plot_grid(plots[[1]], plots[[2]], nrow = 2)
accuracy(plotting.forecast, seis.h.test)

plotting.model <- models[["fit412010"]]
plotting.forecast <- models[["forecast412010"]]
checkresiduals(plotting.model)
residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
plot_grid(residuals.acf, residuals.pacf, nrow=2)
plot_forecast(plotting.forecast) + 
  autolayer(seis.h.test, series="Original") 
plots <- show_overlap_and_diff(
  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
  base.desc = "original series",
  compare.desc = "reconstructed series",
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="% over ts span",
  diff.ylab.sec.accuracy = .01
)
plot_grid(plots[[1]], plots[[2]], nrow = 2)
accuracy(plotting.forecast, seis.h.test)



plotting.model <- models[["fit500010"]]
plotting.forecast <- models[["forecast500010"]]
checkresiduals(plotting.model)
residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
plot_grid(residuals.acf, residuals.pacf, nrow=2)
plot_forecast(plotting.forecast) + 
  autolayer(seis.h.test, series="Original") 
plots <- show_overlap_and_diff(
  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
  base.desc = "original series",
  compare.desc = "reconstructed series",
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="% over ts span",
  diff.ylab.sec.accuracy = .01
)
plot_grid(plots[[1]], plots[[2]], nrow = 2)
accuracy(plotting.forecast, seis.h.test)



for (model in to_test) {
  # model <- c(4,1,2,0,1,0)
  fit.name <- paste0("fit", paste0(model, collapse = ""))
  forecast.name <- paste0("forecast", paste0(model, collapse = ""))
  print("SARIMA(" %s+% paste0(model[1:3], collapse = ",") %s+% ")(" %s+% paste0(model[4:6], collapse = ",") %s+% ")[" %s+% frequency(seis.h.train) %s+% "]")
  if ((fit.name %in% names(models))) {  if ((forecast.name %in% names(models))) {
    plotting.model <- models[[fit.name]]
    plotting.forecast <- models[[forecast.name]]
    checkresiduals(plotting.model)
    residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
    residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
    print(plot_grid(residuals.acf, residuals.pacf, nrow=2))
    print(plot_forecast(plotting.forecast) + 
      autolayer(seis.h.test, series="Original") )
    plots <- show_overlap_and_diff(
      seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
      base.desc = "original series",
      compare.desc = paste0("refit with ", plotting.forecast$method),
      overlap.ylab = "Mean hourly movement (nm)",
      diff.title = "Difference between ts",
      diff.ylab = "nm", diff.ylab.sec="% over ts span",
      diff.ylab.sec.accuracy = .01
    )
    print(plot_grid(plots[[1]], plots[[2]], nrow = 2))
    accuracy(plotting.forecast, seis.h.test)
  }}
}
gg_qqplot(models[["fit500010"]]$residuals)
```


```{r}
#fit <- auto.arima(seis.h.train, d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10, trace = T, parallel = T)
```
 # D=1 m=24
 Fitting models using approximations to speed things up...

 ARIMA(2,0,2)(1,1,1)[24] with drift         : Inf
 ARIMA(0,0,0)(0,1,0)[24] with drift         : 111462.7
 ARIMA(1,0,0)(1,1,0)[24] with drift         : 101954.4
 ARIMA(0,0,1)(0,1,1)[24] with drift         : Inf
 ARIMA(0,0,0)(0,1,0)[24]                    : 111460.9
 ARIMA(1,0,0)(0,1,0)[24] with drift         : 102426.5
 ARIMA(1,0,0)(2,1,0)[24] with drift         : 101379
 ARIMA(1,0,0)(2,1,1)[24] with drift         : Inf
 ARIMA(1,0,0)(1,1,1)[24] with drift         : Inf
 ARIMA(0,0,0)(2,1,0)[24] with drift         : 110647.6
 ARIMA(2,0,0)(2,1,0)[24] with drift         : 101332.9
 ARIMA(2,0,0)(1,1,0)[24] with drift         : 101878.7
 ARIMA(2,0,0)(2,1,1)[24] with drift         : Inf
 ARIMA(2,0,0)(1,1,1)[24] with drift         : Inf
 ARIMA(3,0,0)(2,1,0)[24] with drift         : 101317.9
 ARIMA(3,0,0)(1,1,0)[24] with drift         : 101858.5
 ARIMA(3,0,0)(2,1,1)[24] with drift         : Inf
 ARIMA(3,0,0)(1,1,1)[24] with drift         : Inf
 ARIMA(4,0,0)(2,1,0)[24] with drift         : 101302.7
 ARIMA(4,0,0)(1,1,0)[24] with drift         : 101845
 ARIMA(4,0,0)(2,1,1)[24] with drift         : Inf
 ARIMA(4,0,0)(1,1,1)[24] with drift         : Inf
 ARIMA(5,0,0)(2,1,0)[24] with drift         : 101305.4
 ARIMA(4,0,1)(2,1,0)[24] with drift         : 101304.5
 ARIMA(3,0,1)(2,1,0)[24] with drift         : 101307.7
 ARIMA(5,0,1)(2,1,0)[24] with drift         : 101307.7
 ARIMA(4,0,0)(2,1,0)[24]                    : 101300.7
 ARIMA(4,0,0)(1,1,0)[24]                    : 101843
 ARIMA(4,0,0)(2,1,1)[24]                    : Inf
 ARIMA(4,0,0)(1,1,1)[24]                    : Inf
 ARIMA(3,0,0)(2,1,0)[24]                    : 101316
 ARIMA(5,0,0)(2,1,0)[24]                    : 101303.4
 ARIMA(4,0,1)(2,1,0)[24]                    : 101302.4
 ARIMA(3,0,1)(2,1,0)[24]                    : 101305.8
 ARIMA(5,0,1)(2,1,0)[24]                    : 101305.7

 Now re-fitting the best model(s) without approximations...

 ARIMA(4,0,0)(2,1,0)[24]                    : 101586.6

 Best model: ARIMA(4,0,0)(2,1,0)[24]   
 
 # d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10 m=168
 Fitting models using approximations to speed things up...

 ARIMA(2,1,2)(1,1,1)[168]                    : Inf
 ARIMA(0,1,0)(0,1,0)[168]                    : 122379.8
 ARIMA(1,1,0)(1,1,0)[168]                    : Inf
 ARIMA(0,1,1)(0,1,1)[168]                    : Inf
 ARIMA(0,1,0)(1,1,0)[168]                    : Inf
 ARIMA(0,1,0)(0,1,1)[168]                    : Inf
 ARIMA(0,1,0)(1,1,1)[168]                    : Inf
 ARIMA(1,1,0)(0,1,0)[168]                    : 121638
 ARIMA(1,1,0)(0,1,1)[168]                    : Inf
 ARIMA(1,1,0)(1,1,1)[168]                    : Inf
 ARIMA(2,1,0)(0,1,0)[168]                    : 121468.5
 ARIMA(2,1,0)(1,1,0)[168]                    : Inf
 ARIMA(2,1,0)(0,1,1)[168]                    : Inf
 ARIMA(2,1,0)(1,1,1)[168]                    : Inf
 ARIMA(3,1,0)(0,1,0)[168]                    : 121452.9
 ARIMA(3,1,0)(1,1,0)[168]                    : Inf
 ARIMA(3,1,0)(0,1,1)[168]                    : Inf
 ARIMA(3,1,0)(1,1,1)[168]                    : Inf
 ARIMA(4,1,0)(0,1,0)[168]                    : 121403.9
 ARIMA(4,1,0)(1,1,0)[168]                    : Inf
 ARIMA(4,1,0)(0,1,1)[168]                    : Inf
 ARIMA(4,1,0)(1,1,1)[168]                    : Inf
 ARIMA(5,1,0)(0,1,0)[168]                    : 121361.7
 ARIMA(5,1,0)(1,1,0)[168]                    : Inf
 ARIMA(5,1,0)(0,1,1)[168]                    : Inf
 ARIMA(5,1,0)(1,1,1)[168]                    : Inf
 ARIMA(5,1,1)(0,1,0)[168]                    : 118580.2
 ARIMA(5,1,1)(1,1,0)[168]                    : Inf
 ARIMA(5,1,1)(0,1,1)[168]                    : Inf
 ARIMA(5,1,1)(1,1,1)[168]                    : Inf
 ARIMA(4,1,1)(0,1,0)[168]                    : 118580.3
 ARIMA(5,1,2)(0,1,0)[168]                    : 118581.5
 ARIMA(4,1,2)(0,1,0)[168]                    : 118553.1
 ARIMA(4,1,2)(1,1,0)[168]                    : Inf
 ARIMA(4,1,2)(0,1,1)[168]                    : Inf
 ARIMA(4,1,2)(1,1,1)[168]                    : Inf
 ARIMA(3,1,2)(0,1,0)[168]                    : 118642.8
 ARIMA(3,1,1)(0,1,0)[168]                    : 119155

 Now re-fitting the best model(s) without approximations...

 ARIMA(4,1,2)(0,1,0)[168]                    : 122838.5

 Best model: ARIMA(4,1,2)(0,1,0)[168]               
 
```{r eval=FALSE, include=FALSE}

autoplot(acf(seis.h.train.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
autoplot(pacf(seis.h.train.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()

gg_qqplot(fit$residuals)
gghistogram(fit$residuals)
Box.test(fit$residuals, type="Box-Pierce")
fit.forec <- forecast(fit, h=24*10)
plot(fit.forec, xlim=c(250, 300))


frequency(seis.h.train)
p<-168;d<-0;q<-0;P<-0;D<-1;Q<-0;
paste0("ARIMA(", p, ",", d, ",", q, ")(", P, ",", D, ",", Q, ")_", frequency(seis.h.train))
fit <- Arima(seis.h.train, order=c(p,d,q), seasonal=c(P,D,Q))
fit
#Diagnostics
tsdisplay(residuals(fit), lag.max=168*2)

p<-1;d<-0;q<-2;P<-1;D<-0;Q<-2;
fit <- Arima(seis.h.train, order=c(p,d,q), seasonal=c(P,D,Q))
#Diagnostics
tsdisplay(residuals(fit))
forecast()


```


## MSTL
```{r mstl on msts}
seis.h.train.m <- msts(seis.h.train, c(24, 168))
seis.mstl <- seis.h.train.m %>% mstl(s.window="periodic") #c(15, 7))
seis.mstl.fc <- forecast(seis.mstl, h=168*18)

autoplot(seis.mstl) + xlab("Week")
plot(seis.mstl[,2], main="trend")
lines(trend.s, col="red")
plot(rowSums(seis.mstl[,3:4]), main="seasonality",type="l")
lines(seasonality, col="red")
plot(seis.mstl[,5],  main="residuals")
lines(residuals, col="red")


plotting.model <- seis.mstl
plotting.forecast <- seis.mstl.fc
checkresiduals(plotting.model[,5])
Box.test(plotting.forecast$model$residuals, type="L")
residuals.acf <- autoplot(acf(plotting.model[,5], na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
residuals.pacf <- autoplot(pacf(plotting.model[,5], na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
plot_grid(residuals.acf, residuals.pacf, nrow=2)

plot_forecast(plotting.forecast) + 
  autolayer(seis.h.test, series="Original") +
  xlim(52, 70) + ylim(0, 8750)
plot_forecast(plotting.forecast) + 
  autolayer(seis.h.test, series="Original")
plots <- show_overlap_and_diff(
  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
  base.desc = "train series",
  compare.desc = paste0("refit with ", plotting.forecast$method),
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="% over ts span",
  diff.ylab.sec.accuracy = .01
)
plot_grid(plots[[1]], plots[[2]], nrow = 2)
plots <- show_overlap_and_diff(
  seis.h.test[1:length(plotting.forecast$mean)], plotting.forecast$mean, xlab=seasonality.xlab,
  base.desc = "test series",
  compare.desc = paste0("refit with ", plotting.forecast$method),
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="% over ts span",
  diff.ylab.sec.accuracy = .01
)
plot_grid(plots[[1]], plots[[2]], nrow = 2)
accuracy(plotting.forecast, seis.h.test)
gg_qqplot(plotting.model[,5])

bptests <- data.frame(lag=integer(), `p-value`=numeric())
for (i in 1:10){
  bptests[nrow(bptests) + 1, ] <- c(i, Box.test(plotting.model[,5], lag = i, fitdf = 0, type = "Box-Pierce")$p.value)
}
bptests.plot <- 
  ggplot(bptests, aes(lag, `p.value`), pch="o") + 
    geom_point() + ylim(0,1) +
    labs(title="Box-Pierce tests for lags 1 to 10") +
    geom_hline(yintercept=0.05, linetype="dotted", color = "blue") +
    scale_x_continuous(breaks=1:10)
bptests.plot

```



```{r save RData, include=FALSE}
gc()
save.image(proj_file)
```
